\name{}: machine learning over interaction networks and ontology-based background knowledge predicts drug--target interactions

### 2021-07-12 paper/main.tex
.

Another view on these two approaches is as direct and indirect ways to
predict drug--target interactions. On one hand, molecular information
can be used to directly determine whether two molecules (such as a
drug and protein) have the ability to interact, whereas information
about phenotypic consequences of a drug (drug effects) or disruption
of a protein function can be used to indirectly suggest candidate
drug--target interactions. Molecular features will be specific to a
drug--target pair and we would not expect this information to
propagate through a protein--protein interaction network; the main
information about drug--target interactions that could be obtained
from interactions between proteins is information about binding sites
between proteins that may also be used by a drug molecule (i.e.,
information that protein $P_1$ binds to protein $P_2$ reveals
information about the molecular structures of both $P_1$ and
$P_2$). On the other hand, phenotypic consequences of changes in
protein function, or drug effects, are often a result of aberrant
pathway or network activity and involve more than one protein;
consequently, we expect these features to benefit more from including
information about protein--protein interactions. Moreover, as the
protein--protein interaction relation is not transitive (protein $P_1$
interacting with $P_2$, and $P_2$ relating with $P_3$ does not imply
$P_1$ interacting with $P_3$), we mainly transfer information to the
direct neighborhood of each protein within the PPI graph. Our results
(Table \ref{tab:Results}) confirm the first hypothesis and demonstrate
that molecular features do not benefit from including the interaction
network whereas the indirect, top-down features benefit from the
propagation over the interaction network. While the graph we chose is
based on interactions between proteins, other types of entities and
interactions can also be chosen, in particular similarity networks
\cite{Gottlieb2011}; in such networks, information between nodes
may be transmitted differently than in \name{}.

>>>>

Another view on these two approaches is as direct and indirect ways to
predict drug--target interactions. On one hand, molecular information
can be used to directly determine whether two molecules (such as a
drug and protein) have the ability to interact, whereas information
about phenotypic consequences of a drug (drug effects) or disruption
of a protein function can be used to indirectly suggest candidate
drug--target interactions. Molecular features will be specific to a
drug--target pair and we would not expect this information to
propagate through a protein--protein interaction network; the main
information about drug--target interactions that could be obtained
from interactions between proteins is information about binding sites
between proteins that may also be used by a drug molecule (i.e.,
information that protein $P_1$ binds to protein $P_2$ reveals
information about the molecular structures of both $P_1$ and
$P_2$). On the other hand, phenotypic consequences of changes in
protein function, or drug effects, are often a result of aberrant
pathway or network activity and involve more than one protein;
consequently, we expect these features to benefit more from including
information about protein--protein interactions. Moreover, as the
protein--protein interaction relation is not transitive (protein $P_1$
interacting with $P_2$, and $P_2$ relating with $P_3$ does not imply
$P_1$ interacting with $P_3$), we mainly transfer information to the
direct neighborhood of each protein within the PPI graph. Our results
(Table \ref{tab:Results}) confirm the first hypothesis and demonstrate
that molecular features do not benefit from including the interaction
network whereas the indirect, top-down features benefit from the
propagation over the interaction network. While the graph we chose is
based on interactions between proteins and our results hold true for
such a graph, other types of entities and interactions can also be
chosen, in particular similarity networks \cite{Gottlieb2011}; in such
networks, information between nodes may be transmitted differently
than in \name{}.
